### 2023-06-21 ms.qmd
Updates

The subjects in our experiments were juveniles from two species of skinks, the common garden skink (*L. guichenoti*) and the rainbow skink or delicate skink (*Lampropholis delicata*). These are small (∼35--55 mm snout-vent length (SVL)), oviparous, and generalist skinks that usually share the same habitat in suburban areas throughout south-eastern Australia (Chapple et al. 2011). They also have similar breeding periods, but with some differences in reproductive output: while *L. delicata* lays 1 to 6 eggs in only one clutch per season, *L. guichenoti* clutches are smaller (1-5 eggs per clutch) but they usually lay two clutches per season (Chapple et al. 2011; 2014). This difference in reproductive output through the breeding season may involve different degree of sensitivity to maternal effects and its interaction with other aspects of early environment.

@@

*Eggs collection and incubation* -- Between mid-October 2022 to the end of February 2023, we provided females with a place to lay the eggs by means of small boxes (12.5 L x 8.3 W x 5 H cm) with moist vermiculite inside, that were placed in one extreme of the communal enclosures (see above). We checked for the presence of eggs in the boxes three days a week. After collection, we measured length and width of eggs with a digital calliper to the nearest 0.1 mm and weight them with a (OHAUS, Model spx123) digital scale ± 0.001g error. Then eggs were treated with CORT or vehicle (see CORT and Temperature manipulation below) and were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite). The cups were covered with cling wrap to retain moisture and left in LATWIT 2X5D-R1160 incubators at two different temperatures (see CORT and Temperature manipulation below) until hatching.

*Hatchlings* -- Eggs in the incubator were checked three times a week for hatchlings. After hatchling, we measured juveniles\' SVL and Tail Length (TL) with a rule to the nearest mm and weighted them with a (OHAUS, Model spx123) digital scale ± 0.001g error. We then placed hatchlings in individual enclosures (18.7L x 13.2W x 6.3H cm) and provided them with non-stick matting and a small water dish. During this period, they were sprayed water every day and received 3-6 small *A. domestica* crickets three times a week. All care otherwise follows similar protocols to adults (see above).

@@

To test empirically the effect of early environment we manipulated CORT concentration in eggs and incubated them under one of two temperature regimes (\'Cold\' -- 23ºC ± 3ºC or \'Hot\' -- 30ºC ± 3ºC) in a 2x2 factorial design (Fig. 1a). We first allocated eggs to one of two different treatments: in one of the treatments, eggs were topically supplied with 5µL of CORT dissolved in 70% Ethanol and 30% DMSO (vehicle) at a final (10 pg CORT/mL) concentration (CORT treatment); while eggs in the other group (Control treatment) received an equal volume of the vehicle. CORT concentration employed in the CORT treatment represents 2 standard deviations above the mean natural concentration obtained in eggs from both species (non-published data). Then, eggs were incubated in one of the two previously mentioned temperature regimes (\'Cold\' or \'Hot\') until hatching. The number of eggs per clutch assigned to each hormone and temperature treatment were counterbalanced in both species.

@@

To estimate learning skills, we tested skinks\' ability to locate a food reward in a series of behavioural tasks (Fig. 2) (see Leal and Powell, 2012; Clark *et al.*, 2013). First, we performed a training phase where lizards had to learn to eat from white 3D-printed PLA ramps (9 L x 4 W x 5 H cm) identical to the ones from the experiment except for the colour (see below). We divided this training phase into three stages (see Fig. 2): in the first stage, lizards had to eat a small, frozen cricket (*A. domestica*) from an opaque petri dish (3D x 1.6H cm) placed in the middle of their enclosure; in the second stage, the petri dish with the cricket was placed on top of the white 3D printed ramps; and finally, the cricket was left inside a well (3D x 1.75H cm) on the top of the ramp in the third and last stage. Every trial began when we left the feeding block (petri dish, ramp, or both) inside the enclosure and finished one hour later when we took it away. At the end of the experiment, we recorded whether the cricket has been consumed or not. Trial was considered successful if the lizard could locate and consume the reward, while completion of each stage required the lizards to eat the crickets in at least 5 out of 6 trials. This phase lasted 38 days until all the lizards were able to eat from the ramp; only in one case we decided not to use the lizard because its behaviour was not consistent over the course of the training phase (see Supplementary material).

>>>>

The subjects in our experiments were juveniles from two species of skinks, the common garden skink (*Lampropholis guichenoti*) and the rainbow skink or delicate skink (*L. delicata*). These are small (∼35--55 mm snout-vent length (SVL)), oviparous, and generalist skinks that usually share the same habitat in suburban areas throughout south-eastern Australia ([@chapple2011]). They also have similar breeding periods, but with some differences in reproductive output: while *L. delicata* lays 1 to 6 eggs in only one clutch per season, *L. guichenoti* clutches are smaller (1-5 eggs per clutch) but they usually lay two clutches per season ([@chapple2011]; 2014). This difference in reproductive output through the breeding season may involve different degree of sensitivity to maternal effects and its interaction with other aspects of early environment.

@@

*Eggs collection and incubation* -- Between mid-October 2022 to the end of February 2023, we provided females with a place to lay the eggs by means of small boxes (12.5 L x 8.3 W x 5 H cm) with moist vermiculite inside, that were placed in one extreme of the communal enclosures (see above). We checked for the presence of eggs in the boxes three days a week. After collection, we measured length and width of eggs with a digital caliper to the nearest 0.1 mm and weight them with a (OHAUS, Model spx123) digital scale ± 0.001g error. Then eggs were treated with CORT or vehicle (see CORT and Temperature manipulation below) and were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite). The cups were covered with cling wrap to retain moisture and left in LATWIT 2X5D-R1160 incubators at two different temperatures (see CORT and Temperature manipulation below) until hatching.

*Hatchlings* -- Eggs in the incubator were checked three times a week for hatchlings. After hatchling, we measured juveniles' SVL and Tail Length (TL) with a rule to the nearest mm and weighted them with a (OHAUS, Model spx123) digital scale ± 0.001g error. We then placed hatchlings in individual enclosures (18.7L x 13.2W x 6.3H cm) and provided them with non-stick matting and a small water dish. During this period, they were sprayed water every day and received 3-6 small *A. domestica* crickets three times a week. All care otherwise follows similar protocols to adults (see above).

@@

To test empirically the effect of early environment we manipulated CORT concentration in eggs and incubated them under one of two temperature regimes ('Cold' -- 23ºC ± 3ºC or 'Hot' -- 30ºC ± 3ºC) in a 2x2 factorial design (Fig. 1a). We first allocated eggs to one of two different treatments: in one of the treatments, eggs were topically supplied with 5µL of CORT dissolved in 70% Ethanol and 30% DMSO (vehicle) at a final (10 pg CORT/mL) concentration (CORT treatment); while eggs in the other group (Control treatment) received an equal volume of the vehicle. CORT concentration employed in the CORT treatment represents 2 standard deviations above the mean natural concentration obtained in eggs from both species (non-published data). Then, eggs were incubated in one of the two previously mentioned temperature regimes ('Cold' or 'Hot') until hatching. The number of eggs per clutch assigned to each hormone and temperature treatment were counterbalanced in both species.

@@

To estimate learning skills, we tested skinks' ability to locate a food reward in a series of behavioural tasks (Fig. 2) (see Leal and Powell, 2012; Clark *et al.*, 2013). First, we performed a training phase where lizards had to learn to eat from white 3D-printed PLA ramps (9 L x 4 W x 5 H cm) identical to the ones from the experiment except for the colour (see below). We divided this training phase into three stages (see Fig. 2): in the first stage, lizards had to eat a small, frozen cricket (*A. domestica*) from an opaque petri dish (3D x 1.6H cm) placed in the middle of their enclosure; in the second stage, the petri dish with the cricket was placed on top of the white 3D printed ramps; and finally, the cricket was left inside a well (3D x 1.75H cm) on the top of the ramp in the third and last stage. Every trial began when we left the feeding block (petri dish, ramp, or both) inside the enclosure and finished one hour later when we took it away. At the end of the experiment, we recorded whether the cricket has been consumed or not. Trial was considered successful if the lizard could locate and consume the reward, while completion of each stage required the lizards to eat the crickets in at least 5 out of 6 trials. This phase lasted 38 days until all the lizards were able to eat from the ramp; only in one case we decided not to use the lizard because its behaviour was not consistent over the course of the training phase (see Supplementary material).
